\documentclass[lscape,PhD]{msu-thesis}

\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[colorlinks=True,linkcolor=black,urlcolor=blue,citecolor=black]{hyperref}       % hyperlinks
\usepackage{xurl}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{textgreek}
\usepackage{bm}
\usepackage{units}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{xcolor}
\usepackage{tabularx,booktabs} % pretty captions and table formatting
\usepackage{gensymb}        % Litterally needed just for degree symbol
\usepackage{mathtools}
\usepackage{xspace}         % Helpful for formatting fancy math text
\usepackage{graphicx}
\usepackage{multirow}       % merge cells in table
% \usepackage{afterpage}      % Page breaking for single tables
\usepackage{pdflscape}      % Page manipulate like landscape/portrait
\usepackage{arydshln}       % Tables with dashed lines
\usepackage{newtxtext,newtxmath}
\usepackage[justification=centering]{caption}
\usepackage[capitalize]{cleveref}
\usepackage[style=numeric,sorting=none]{biblatex}
\usepackage{listings}
\usepackage{makecell}
\usepackage{dark_matter}    % User defined macros

% \setlength{\afterchapskip}{\onelineskip}
\addbibresource{refs.bib}

\title{Leveraging Multi-Messenger Astrophysics for Dark Matter Searches}
\author{Daniel Nicolas Salazar-Gallegos}
\dualmajor{Physics and Astronomy}{Computational Mathematics, Science, and Engineering}
\date{2024}

\begin{document}
\frontmatter
\maketitlepage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

    This dissertation enhances the search for dark matter (DM) through a multi-wavelength and multi-messenger approach by combining data from leading gamma-ray and neutrino observatories.
    The first analysis combines data across five gamma-ray observatories, including the High Altitude Water Cherenkov (HAWC) Observatory and four others.
    This foundational study probed the largest DM mass space that spans from 5 GeV up to 100 TeV.
    This collaboration  achieved a threefold improvement in sensitivity to DM annihilation from dwarf galaxies.
    The collaborative framework developed among these five experiments also set a goal for including neutrino observatories in DM searches.
    Though no significant dark matter signals were detected, more stringent upper limits were set on the velocity-weighted annihilation cross-section, \sv.

    A pivotal aspect of the research involved developing new analytical methods within the IceCube Neutrino Observatory that achieved an order of magnitude increase in sensitivity for heavy dark matter annihilation.
    These improvements set the stage for multi-messenger dark matter searches with IceCube and gamma-ray observations.

    Within HAWC, computational methods were optimized to speed up the analysis pipeline by an order of magnitude.
    HAWC is therefore able to take on more ambitious DM searches and observe many more potential DM sources.

    The combined efforts from HAWC and IceCube form the basis for a novel multi-messenger strategy to probe dark matter annihilation searches.
    The research culminates in a preliminary mock limit from HAWC data and IceCube simulation.
    These analyses showcase the enhanced capability of detecting dark matter in the mass region above 1 PeV.
    With these two instruments alone, we could see two to threefold improvement to our sensitivity to heavy dark matter compared to these observatories searching for DM independently.
    This dissertation presents a comprehensive multi-instrument and multi-wavelength approach that provides a robust framework for future dark matter research.
    This work will be the foundation for a multi-messenger approach in the pursuit of unveiling the universe's dark sector.

\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage
\pagestyle{plain}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{Acknowledgments}
\DoubleSpacing

This dissertation was ambitious and benefited greatly from support and love of my friends, family, and mentors.
These people believed in me at times when I found it hard to believe in myself.
Thank you so much, all of you, for encouraging me to follow my ambition and achieve great things.
Several of you come to mind and below I thank you individually in no particular order.

\textbf{My mentors}

Ben Hooberman, my undergraduate mentor who took a risk on an inexperienced undergraduate researcher.
He gave me the opportunities in physics I was craving and exposed me to the world of research.
The first of many doors opened when I started with Ben, and I am tremendously grateful.

Kirsten Tollefson, my graduate advisor and chair to my thesis committee.
She guided my ambitious program and kept me grounded.
Her direction and training taught me a lot about successful research and building strong, collaborative teams.
I will miss the days of gaming out research and project plans.
Thanks a ton for making graduate school a pleasant experience for me.

Pat Harding, member on my committee and mentor at Los Alamos National Lab.
I learned ways of thinking pragmatically towards solving problems with Pat.
He helped me lower the stakes and focused on the core goals in research.
Thanks for taking a chance on me and introducing me to Los Alamos, truly an enchanting place.

Mehr Nisa, previously a MSU postdoc, now faculty.
We arrived at MSU at about the same time and was the only postdoc/peer who was also working on both HAWC and IceCube.
Her expertise and guidance was critical in developing the ambitious research of this dissertation.
It's also been inspiring and a source of community to grow in our careers in parallel.
Thanks a ton for the mentorship, and supporting my many questions.

Jim Linnemann, faculty at MSU in HAWC and rigorous physics mentor.
Jim has been a role model of scientific rigor who has guided my physics skill set since arriving at MSU.
He is among the most exposed to my research and has similarly enabled my ambition.
It's been a pleasure to work with Jim and I thank you tremendously for your input to my research.

Thank you to Kendall Mahn, who saw my potential from the moment of application to MSU.
You cultivated the fight in me that's been exercised in leadership positions at MSU.

Thank you to the remaining members of my thesis committee: Claudio Kopper, Brian O'Shae, and Filomena Nunes.
You have been a great committee providing constructive feedback on this dissertation.

Thank you to the MSU postdocs Hans Niederhausen, Brian Clark, and Rob Halliday who provided substantial intellectual training and were great people to work with.

Thank you to the Los Alamos HAWC cast, Andrea Albert, Brenda Dingus, Kelly Malone and Mora Durocher for holding space for me and creating a second research home.
My experience at LANL was invaluable and left a positive, indelible mark on my experience as a grad student.

Thank you to grad students, Allison Peisker, Joe Lundeen, Jessie Micallef, and Minjin Jeong who showed me the ropes on research with HAWC and IceCube.
You provided a strong foundation for me to build on with my research.

\textbf{My Family}

Familia is core principle and social support within the Mexican culture, and I am no exception.
Thank you to my mom, and brothers, Miguel and Chris, especially who have kept me grounded through my Ph.D. and came to my rescue at my lowest points in it.
Thank you to my extended family in Texas and Mexico who have always had a door open for me and supported me as I redefined who I am.

We came from simple means, lives, and beginnings.
I hope that this effort and title inspires those within the family of the great things we are capable of.

\textbf{My Friends}

Thank you to the friends I made at Los Alamos and being the first to accept me as my authentic self.
You were among the first to show me to embrace myself.
Thank you specifically to Eames Bennet and Jeremy Monroe who opened their homes to me and did not ask for repayment.
A place to stay was invaluable for enabling my visits to LANL.

Thank you to the climbing belaytionships that are the backbone to my sanity through the COVID-19 pandemic and grad school.
David Shane, Jon Fleck, and Alex Harnisch, thank y'all for belaying me on so many routes and motivating me to push hard.
The grit and adventures we've had are cherished memories.

Thank you to Tristan Wells.
It's crazy that I met someone so chill and easy to get along with by accident.
I'm so happy to have grown in my Ph.D. with you and would not have chosen another roommate.
Thank you for being along for the ride and being a part of my life for these six years.

Thank you to Jenny and her late child Kaydince.
I only got to know you for a short amount of time, but you made a big splash.
I wish the best for you and that you are successful in your pursuit to cure the blind.

Thank you to my partner Corey Snyder.
You came at the best time to see me under the most stress and loved me nonetheless.
Your support and love propped me up through the roughest moment of my dissertation.
Quite literally, this work would not have been completed without your support.
Thank you for your kindness, motivation, love, laughs, safety, adventure, and thoughtful conversation.

\textbf{Institutions}

Thank you to the National Science Foundation for funding me for three years of my Ph.D.
Thank you to MSU for funding another two through their fellowship program.
Thank you to GEM for their fellowship program in partnership with LANL.
Thank you to LANL for always leaving a desk for me after my first summer in 2019.
Thank you to HAWC and IceCube for providing me with the resources and expertise to thrive.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage

\SingleSpacing
\tableofcontents*
% \DoubleSpacing

\clearpage
\mainmatter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Dark Matter in the Cosmos}\label{sec:dm_cosmos}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{chapters/DarkMatterintheCosmos}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{High Altitude Water Cherenkov (HAWC) Observatory}\label{sec:hawc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{chapters/hawc.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Glory Duck: Multi-wavelength Search for Dark Matter Annihilation Towards Dwarf Spheroidal Galaxies}\label{sec:glory_duck}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{chapters/gloryduck.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Multithreading HAWC analyses for Dark Matter Searches} \label{sec:multithread}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{chapters/multithreading_dm.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{IceCube Neutrino Observatory}\label{sec:ice3}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{chapters/ic3.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Heavy Dark Matter Annihilation Search with IceCube's North Sky Track Data} \label{sec:ic3_dm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{chapters/ice3_darkmatter.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Nu Duck: Conclusions and Future Directions}\label{sec:nu_duck}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{chapters/nu_duck.tex}

\SingleSpacing

\printbibliography

\DoubleSpacing

\begin{appendices}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \chapter{Multi-Experiment Supplementary Figures}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{appendices/A1_gd_spatial_maps.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \chapter{Multithreading Dark Matter Analyses Supplementary Material}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{appendices/A2_mtd_dm_hawc.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \chapter{IceCube Heavy Dark Matter Analysis Supplementary Material}\label{apdx:ic3DM_supp}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{appendices/A3_ic3_dm.tex}

\end{appendices}

\end{document}